---
title: "Lasso Regression"
subtitle: "Learn how to build a Lasso model with tidymodels"
author: "Prof. Dr. Jan Kirenz"
output:
 html_document:
  code_download: true 
  css: style.css 
  fig_height: 6
  fig_width: 8
  highlight: tango
  number_sections: yes
  theme: paper
  toc: yes
  toc_depth: 3
  toc_float: 
    collapsed: false
    smooth_scroll: true 
  df_print: paged
---


```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE, 
                      dpi = 300, 
                      fig.width = 8, 
                      fig.height = 5)
library(scales)
library(tidyverse)

```


In this tutorial we take a look at the dataset *The Office* to show how to build a lasso regression model and choose regularization parameters.

*Source: Julia Silge, [LASSO regression using tidymodels and #TidyTuesday data for The Office]( https://juliasilge.com/blog/lasso-the-office/).*

</br>

# Explore the data

Our modeling goal here is to predict the IMDB ratings for episodes of *The Office* based on the other characteristics of the episodes in the [#TidyTuesday dataset](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-17/readme.md). 

## Data

There are two datasets, one with the ratings and one with information like director, writer, and which character spoke which line. The episode numbers and titles are not consistent between them, so we can use regular expressions to do a better job of matching the datasets up for joining.

```{r}
library(scales)
library(tidyverse)

ratings_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-17/office_ratings.csv')

remove_regex <- "[:punct:]|[:digit:]|parts |part |the |and"

office_ratings <- ratings_raw %>%
  transmute(
    episode_name = str_to_lower(title),
    episode_name = str_remove_all(episode_name, remove_regex),
    episode_name = str_trim(episode_name),
    imdb_rating
  )

office_info <- schrute::theoffice %>%
  mutate(season = as.numeric(season),
         episode = as.numeric(episode),
         episode_name = str_to_lower(episode_name),
         episode_name = str_remove_all(episode_name, remove_regex),
         episode_name = str_trim(episode_name)) %>%
  select(season, episode, episode_name, director, writer, character) 

office_info
```

We are going to use several different kinds of features for modeling. First, let's find out how many times characters speak per episode.

```{r}
characters <- office_info %>%
  count(episode_name, character) %>%
  add_count(character, wt = n, name = "character_count") %>%
  filter(character_count > 800) %>%
  select(-character_count) %>%
  pivot_wider(names_from = character, 
              values_from = n, 
              values_fill = list(n = 0))

characters
```

Next, let's find which directors and writers are involved in each episode. I'm choosing here to combine this into one category in modeling, for a simpler model, since these are often the same individuals.

```{r}

creators <- office_info %>%
  distinct(episode_name, director, writer) %>%
  pivot_longer(director:writer, names_to = "role", values_to = "person") %>%
  separate_rows(person,  sep = ";") %>%
  add_count(person) %>%
  filter(n > 10) %>%
  distinct(episode_name, person) %>%
  mutate(person_value = 1) %>% 
  pivot_wider(names_from = person, 
              values_from = person_value, 
              values_fill = list(person_value = 0))

creators
```

Next, let's find the season and episode number for each episode, and then finally let's put it all together into one dataset for modeling.

```{r}
office <- office_info %>%
  distinct(season, episode, episode_name) %>%
  inner_join(characters) %>%
  inner_join(creators) %>%
  inner_join(office_ratings %>%
               select(episode_name, imdb_rating)) %>%
  janitor::clean_names()

office
```

## EDA

There are lots of [great examples of EDA on Twitter](https://twitter.com/search?q=%23TidyTuesday); I especially encourage you to check out [the screencast David Robinson](https://youtu.be/_IvAubTDQME), which is similar in spirit to the modeling here but he includes more EDA. 

```{r}

office %>%
  ggplot(aes(episode, imdb_rating, fill = as.factor(episode))) +
  geom_boxplot(show.legend = FALSE)

```

Ratings are higher for episodes later in the season. What else is associated with higher ratings? Let's use lasso regression to find out! 

# Train a model

## Data split

We can start by loading the tidymodels metapackage, and splitting our data into training and testing sets.

```{r}
library(tidymodels)

office_split <- initial_split(office, strata = season)

office_train <- training(office_split)
office_test <- testing(office_split)

```

## Data preprocessing

Then, we build a recipe for data preprocessing.

- First, we must tell the `recipe()` what our model is going to be (using a formula here) and what our training data is.
- Next, we update the role for `episode_name`, since this is a variable we might like to keep around for convenience as an identifier for rows but is not a predictor or outcome.
- Next, we remove any numeric variables that have zero variance.
- As a last step, we normalize (center and scale) the numeric variables. We need to do this because it's important for lasso regularization.

The object `office_rec` is a recipe that has **not** been trained on data yet (for example, the centered and scaling has not been calculated) and `office_prep` is an object that **has** been trained on data. The reason we use `strings_as_factors = FALSE` here is that the ID column `episode_name` is of type character (as opposed to, say, integers).

```{r}
office_rec <- recipe(imdb_rating ~ ., data = office_train) %>%
  update_role(episode_name, new_role = "ID") %>%
  step_zv(all_numeric(), -all_outcomes()) %>%
  step_normalize(all_numeric(), -all_outcomes())

office_prep <- office_rec %>%
  prep(strings_as_factors = FALSE)
```

## Specify and fit model

Now it's time to **specify** and then **fit** our models. 

Here we set up one model specification for lasso regression; I picked a value for `penalty` (sort of randomly) and I set `mixture = 1` for lasso. 

I am using a [`workflow()`](https://tidymodels.github.io/workflows/) in this example for convenience; these are objects that can help you manage modeling pipelines more easily, with pieces that fit together like Lego blocks. 

You can `fit()` a workflow, much like you can fit a model, and then you can pull out the fit object and `tidy()` it!

```{r}

lasso_spec <- linear_reg(penalty = 0.1, mixture = 1) %>%
  set_engine("glmnet")

wf <- workflow() %>%
  add_recipe(office_rec) # here we use office_rec, not office_prep

lasso_fit <- wf %>%
  add_model(lasso_spec) %>%
  fit(data = office_train)

lasso_fit %>%
  pull_workflow_fit() %>%
  tidy()

```

If you have used glmnet before, this is the familiar output where we can see (here, for the most regularized examples) what contributes to higher IMDB ratings.

# Tune lasso parameters

So we fit one lasso model, but how do we know the right regularization parameter `penalty`? We can figure that out using resampling and tuning the model. 

Let's build a set of bootstrap resamples, and set `penalty = tune()` instead of to a single value. We can use a function `penalty()` to set up an appropriate grid for this kind of regularization model.


```{r}
set.seed(1234)

office_boot <- bootstraps(office_train, 
                          strata = season)

tune_spec <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

lambda_grid <- grid_regular(penalty(), levels = 50)

```

Now it's time to tune the grid, using our workflow object.

# Grid tuning

```{r}
doParallel::registerDoParallel()

set.seed(2020)

lasso_grid <- 
  tune_grid(
  wf %>% 
    add_model(tune_spec),
    resamples = office_boot,
    grid = lambda_grid
)
```

What results did we get?

```{r}
lasso_grid %>%
  collect_metrics()
```

Let's take a look at a visualization of performance with the regularization parameter.

## Performance plot

```{r fig.width = 8, fig.height = 7}

lasso_grid %>%
  collect_metrics() %>%
  ggplot(aes(penalty, mean, color = .metric)) +
  geom_errorbar(aes(ymin = mean - std_err,
                    ymax = mean + std_err),
                alpha = 0.5) +
  geom_line(size = 1.5) +
  facet_wrap(~.metric, scales = "free", nrow = 2) +
  scale_x_log10() +
  theme(legend.position = "none")
```

This is a great way to see that regularization helps this modeling a lot. We have a couple of options for choosing our final parameter, such as 

- `select_by_pct_loss()` or 
- `select_by_one_std_err()`, 

but for now let's stick with just picking the lowest RMSE. After we have that parameter, we can finalize our workflow, i.e. update it with this value.

## Finalize model

```{r}

lowest_rmse <- lasso_grid %>%
  select_best("rmse")

final_lasso <- 
  finalize_workflow(
    wf %>% 
      add_model(tune_spec),
      lowest_rmse)

```

## Most important variables

We can then fit this finalized workflow on our training data. While we're at it, let's see what the most important variables are using the [vip](https://koalaverse.github.io/vip/) package.

```{r}
library(vip)

final_lasso %>%
  fit(office_train) %>%
  pull_workflow_fit() %>%
  vi(lambda = lowest_rmse$penalty) %>%
  mutate(Importance = abs(Importance),
         Variable = fct_reorder(Variable, Importance)) %>%
  ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
  geom_col() +
  scale_x_continuous(expand = c(0,0)) +
  labs(y = NULL)

```

## last_fit()

And then, finally, let's return to our test data. The tune package has a function `last_fit()` which is nice for situations when you have tuned and finalized a model or workflow and want to fit it one last time on your training data and evaluate it on your testing data. You only have to pass this function your finalized model/workflow and your split.

```{r}
last_fit(final_lasso,
         office_split) %>% # note that we use office_split
  collect_metrics()
```
